### 2019-04-09 papers/journal/encoding.tex
Hyphen problem fixed

\subsubsection{Constraints on presence, activity of the molecule, and pairing matrix}
%
We need the following constraints \texttt{EdgeC},
\texttt{ActivityC} and \texttt{PairingC} to encode the basic structure of VTSs.
%
For an edge to exist it should have one molecule present. 
%
\begin{align*}
  \texttt{PresentE} = \bigwedge\limits_{i,j,q} (\bigvee_m e_{i,j,q,m} )\limplies e_{i,j,q}
\end{align*}
The edge labels are subset of the node labels of the source and target nodes.
\begin{align*}
  \texttt{LabelE} = \bigwedge\limits_{i,j,q,m} e_{i,j,q,m} \limplies (n_{i,m} \land n_{j,m} )
\end{align*}
Self edges are not allowed. 
\begin{align*}
  \texttt{SelfE} = \bigwedge\limits_{i,q} \neg e_{i,i,q}
\end{align*}
\noindent We define \texttt{EdgeC} to collect all the edge related constraints. 
%
\texttt{EdgeC} states that each edge has at least one molecule,
 edge labels are consistent with node labels and there are no self-loops.    
\begin{align*}
  \texttt{EdgeC} = \texttt{PresentE} \land \texttt{LabelE}  \land \texttt{SelfE} 
\end{align*}
\noindent If a molecule is active on an edge, it should be present on the edge.
%
\begin{align*}
  \texttt{ActiveMolE} = \bigwedge\limits_{i,j,q,m} b_{i,j,q,m} \limplies e_{i,j,q,m}\label{eq:f1}
\end{align*}
A molecule should be present to be active on a node.  
\begin{align*}
    \texttt{ActiveMolN} = \bigwedge\limits_{i,m} a_{i,m} \limplies n_{i,m}
\end{align*}
\noindent We define \texttt{ActivityC} to group all the molecule activity related constraints. 
\texttt{ActivityC} states that all active molecules are present.
\begin{align*}
  \texttt{ActivityC} = \texttt{ActiveMolE} \land \texttt{ActiveMolN} 
\end{align*}
\noindent We divide the pairing matrix into Q-SNAREs and R-SNAREs blocks and set the dependency bits of R-SNAREs to R-SNAREs in the pairing matrix to be all 0's.
%It reduces the search space.
% Condition on $p_{kk'}$. \ashu{why???} .
%\begin{align}
%  \bigwedge\limits_{(x < M/2 \, \land  \, y < M/2) \lor  (x >= M/2 \land y >= M/2)} \neg p(x,y)
%  \tag{V6}\label{eq:c3}
%\end{align}

@@

\subsubsection{Well-fused constraints}
Constraint \texttt{Fusion1} encodes that each edge must fuse with
its destination node.
%
Constraint \texttt{Fusion2} encodes that each edge should not
be able to potentially fuse with any node other than its destination node.
\begin{align*}
\texttt{Fusion1} =\;&
\bigwedge\limits_{i,j,q} e_{i,j,q} \limplies
\bigvee_{{m_1,m_2,m_3,m_4}} (\Land_{l=1}^4 ( b_{i,j,q,m_l} \lor a_{j,m_l} ) \land 
\Lor\limits_{l=1}^4 b_{i,j,q,m_l} \land\\
& \qquad \qquad \qquad \qquad \Lor\limits_{l=1}^4 a_{j,m_l}\land p_{\{m_1,m_2,m_3,m_4\}})
\\
\texttt{Fusion2} =\;&
\bigwedge\limits_{\mathclap{i,j,q,m_1,m_2,m_3, l \in \{1,..,3\}}} b_{i,j,q,m_1} \land ..\land b_{i,j,q,m_l} \limplies \\
& \hspace{2cm} \neg 
\bigvee_{\mathclap{j \neq j^{\prime}, m_{l+1}^{\prime},..,m_{4}^{\prime}}} ( a_{j^{\prime},m_{l+1}^{\prime}} \land .. \land a_{j^{\prime},m_4^{\prime}} \land p_{\{m_1,..,m_l,m^{\prime}_{l+1},..,m^{\prime}_4\}})
\\
\end{align*}
We define \texttt{FusionC} to group all the fusion related constraints.
% 
\texttt{FusionC} states that every edge in the graph is well-fused.
\begin{align*}
\texttt{FusionC} = \texttt{Fusion1} \land \texttt{Fusion2} 
\end{align*}
% Fusion rules consist of two different mechanisms.
% \begin{enumerate}
% \item  A \textbf{pairing mechanism} which determines compatible Q-R pairs on vesicles and compartments that can cause fusion.
% \item \textbf{regulatory mechanisms} on the edges and nodes (possibly distinct) which regulates molecules activity based on the presence/absence of other molecules on the corresponding node or edge.
% \end{enumerate}
% Boolean constraint~\eqref{eq:fuse1} and~\eqref{eq:fuse2} are used to encode well-fused
% property.

@@

\subsection{Encoding for $\fbmc$ search problem}
\label{enc:bmc}
\noindent Our first attempt was to encode the existential condition as a $\fbmc$ search problem~\cite{shukla2017discovering}.
%
%In this section we will present the $\fbmc$ encoding of the same.
%The basic idea is to search the solution only upto a certain depth $w$. 
%
%\begin{align*}
%\npath(m, i, j, w) = \bigvee\limits_{l = 1}^{w-1} (\texttt{Length} (m-path_{ij}) == l)
%\end{align*}
%
%
%\begin{align*}
%\bigwedge\limits_{i,j,m} (\bigvee_{q} (\texttt{bvEdge}[i][j][q] \unrhd m) == 1) \limplies \npath(m, j, i, w)
%\end{align*}
%Although the techniques used for $\fbmc$ and $\fsmt$ is very different because we are 
%In $\fbmc$ we only partially unroll the program to check for the solutions within that subset. 
%in the case of $\fbmc$, the encoding is not considerably different.
% 
The main differences between $\fsmt$ and $\fbmc$ are the representation of variables and the encoding of 
%I n comparison to $\fsmt$, the $\fbmc$ search encodes 
reachability (both for the stability and connectivity).  
%]
Here, we will only present the stability encoding.
%the rest of the encodings are similar to the one presented in the previous section, Section~\ref{enc:smt}.

@@

%if all the pairs of the present molecule are present on the same edge than all of them inhibit each other resulting in making all the molecules inactive, we term this process ``pairing inhibition".
%
The constraints for the variants can be given to a SAT or a SMT solver to find
VTSs that belong to the variants.
\subsection{Encoding for $\fqbf$ search problem}\label{enc:univ-cond}
%
\input{satqbf-table}
\noindent The specification for universal condition~\ref{sec:property} requires quantifier alternation. 
%
Most SAT solvers can not efficiently handle the quantifier in the formulas. 
% 
Although, it is possible to encode the universal condition as a SAT or SMT problem.
%
There is always a danger of a blow-up in the formula size, making it harder for the SAT solver to handle. 
%Which may blow up the formula size for the solver to solve.
% 
As an experiment, we have implemented a direct SMT version (encoding not presented) of the universal problem extending the current $\fsmt$ search encodings.
%
The basic idea is to conjunct the formula for each combination value of the quantified variable. 
%
The Table~\ref{tab:satqbf-graph}, shows the search results of this encoding for the universal condition with 4 nodes and 3-connected graphs. 
%
The connectivity constraint and connectedness constraints are tagged with subscript $U$, $\texttt{StabilityC}_{U}$, $\texttt{Connected}(k)_{U}$,  to distinguish from earlier presented encodings of existential condition.
%
The SMT extended version was not able to scale it to the desired results and either timed-out (T/O) or ran out of memory (M/O).
%
This can be due to the fact that directly translating a quantified formula into a quantifier-free by introducing a vector of variables for each boolean possibility can lead to exponential increase in the formula size.
%
%
\subsubsection{$\fqbf$ search encoding of the universal condition}
We will describe several constraints that encode universal condition based on QBF formulation in this section.
%
To avoid cumbersome notation, we will not explicitly write the ranges
of the indexing in the constraints.
%
$i$ and $j$ will range over nodes, i.e., from $1$ to $\nu$.
%
$m$ will range over molecules, i.e., from $1$ to $\mu$.
%
$q$ will range over edges between two nodes, i.e., from $1$ to $\pi$.
%

>>>>

\subsubsection{Constraints on presence, activity of the molecule, and pairing matrix}
%
We need the following constraints \texttt{EdgeC},
\texttt{ActivityC} and \texttt{PairingC} to encode the basic structure of VTSs.
%
For an edge to exist it should have one molecule present. 
%
\begin{align*}
  \texttt{PresentE} = \bigwedge\limits_{i,j,q} (\bigvee_m e_{i,j,q,m} )\limplies e_{i,j,q}
\end{align*}
The edge labels are subset of the node labels of the source and target nodes.
\begin{align*}
  \texttt{LabelE} = \bigwedge\limits_{i,j,q,m} e_{i,j,q,m} \limplies (n_{i,m} \land n_{j,m} )
\end{align*}
Self edges are not allowed. 
\begin{align*}
  \texttt{SelfE} = \bigwedge\limits_{i,q} \neg e_{i,i,q}
\end{align*}
\noindent We define \texttt{EdgeC} to collect all the edge related constraints. 
%
\texttt{EdgeC} states that each edge has at least one molecule,
 edge labels are consistent with node labels and there are no self-loops.    
\begin{align*}
  \texttt{EdgeC} = \texttt{PresentE} \land \texttt{LabelE}  \land \texttt{SelfE} 
\end{align*}
\noindent If a molecule is active on an edge, it should be present on the edge.
%
\begin{align*}
  \texttt{ActiveMolE} = \bigwedge\limits_{i,j,q,m} b_{i,j,q,m} \limplies e_{i,j,q,m}\label{eq:f1}
\end{align*}
A molecule should be present to be active on a node.  
\begin{align*}
    \texttt{ActiveMolN} = \bigwedge\limits_{i,m} a_{i,m} \limplies n_{i,m}
\end{align*}
\noindent We define \texttt{ActivityC} to collect all the molecule activity related constraints. 
\texttt{ActivityC} states that all active molecules are present.
\begin{align*}
  \texttt{ActivityC} = \texttt{ActiveMolE} \land \texttt{ActiveMolN} 
\end{align*}
\noindent We divide the pairing matrix into Q-SNAREs and R-SNAREs blocks and set the dependency bits of R-SNAREs to R-SNAREs in the pairing matrix to be all 0's.
%It reduces the search space.
% Condition on $p_{kk'}$. \ashu{why???} .
%\begin{align}
%  \bigwedge\limits_{(x < M/2 \, \land  \, y < M/2) \lor  (x >= M/2 \land y >= M/2)} \neg p(x,y)
%  \tag{V6}\label{eq:c3}
%\end{align}

@@

\subsubsection{Well-fused constraints}
Constraint \texttt{Fusion1} encodes that each edge must fuse with
its destination node.
%
Constraint \texttt{Fusion2} encodes that each edge should not
be able to potentially fuse with any node other than its destination node.
\begin{align*}
\texttt{Fusion1} =\;&
\bigwedge\limits_{i,j,q} e_{i,j,q} \limplies
\bigvee_{{m_1,m_2,m_3,m_4}} (\Land_{l=1}^4 ( b_{i,j,q,m_l} \lor a_{j,m_l} ) \land 
\Lor\limits_{l=1}^4 b_{i,j,q,m_l} \land\\
& \qquad \qquad \qquad \qquad \Lor\limits_{l=1}^4 a_{j,m_l}\land p_{\{m_1,m_2,m_3,m_4\}})
\\
\texttt{Fusion2} =\;&
\bigwedge\limits_{\mathclap{i,j,q,m_1,m_2,m_3, l \in \{1,..,3\}}} b_{i,j,q,m_1} \land ..\land b_{i,j,q,m_l} \limplies \\
& \hspace{2cm} \neg 
\bigvee_{\mathclap{j \neq j^{\prime}, m_{l+1}^{\prime},..,m_{4}^{\prime}}} ( a_{j^{\prime},m_{l+1}^{\prime}} \land .. \land a_{j^{\prime},m_4^{\prime}} \land p_{\{m_1,..,m_l,m^{\prime}_{l+1},..,m^{\prime}_4\}})
\\
\end{align*}
We define \texttt{FusionC} \change{Use of the word collect looks weird! change it.}to collect all the fusion related constraints.
% 
\texttt{FusionC} states that every edge in the graph is well-fused.
\begin{align*}
\texttt{FusionC} = \texttt{Fusion1} \land \texttt{Fusion2} 
\end{align*}
% Fusion rules consist of two different mechanisms.
% \begin{enumerate}
% \item  A \textbf{pairing mechanism} which determines compatible Q-R pairs on vesicles and compartments that can cause fusion.
% \item \textbf{regulatory mechanisms} on the edges and nodes (possibly distinct) which regulates molecules activity based on the presence/absence of other molecules on the corresponding node or edge.
% \end{enumerate}
% Boolean constraint~\eqref{eq:fuse1} and~\eqref{eq:fuse2} are used to encode well-fused
% property.

@@

\subsection{Encoding for $\fbmc$ search problem}
\label{enc:bmc}
\noindent Our first attempt was to encode the existential condition as a $\fbmc$ search \unsure{Should we cite it here or not?}problem~\cite{shukla2017discovering}.
%
%In this section we will present the $\fbmc$ encoding of the same.
%The basic idea is to search the solution only upto a certain depth $w$. 
%
%\begin{align*}
%\npath(m, i, j, w) = \bigvee\limits_{l = 1}^{w-1} (\texttt{Length} (m-path_{ij}) == l)
%\end{align*}
%
%
%\begin{align*}
%\bigwedge\limits_{i,j,m} (\bigvee_{q} (\texttt{bvEdge}[i][j][q] \unrhd m) == 1) \limplies \npath(m, j, i, w)
%\end{align*}
%Although the techniques used for $\fbmc$ and $\fsmt$ is very different because we are 
%In $\fbmc$ we only partially unroll the program to check for the solutions within that subset. 
%in the case of $\fbmc$, the encoding is not considerably different.
% 
The main differences between $\fsmt$ and $\fbmc$ are the representation of variables and the encoding of 
%I n comparison to $\fsmt$, the $\fbmc$ search encodes 
reachability (both for the stability and connectivity).  
%]
Here, we will only present the stability encoding.
%the rest of the encodings are similar to the one presented in the previous section, Section~\ref{enc:smt}.

@@

%if all the pairs of the present molecule are present on the same edge than all of them inhibit each other resulting in making all the molecules inactive, we term this process ``pairing inhibition".
%
The constraints for the variants can be given to a SAT or a SMT solver to find
VTSs that belong to the variants.
\subsection{Encoding for $\fqbf$ search problem}\label{enc:univ-cond}
%
\input{satqbf-table}
\noindent The specification for universal condition~\ref{sec:property} requires quantifier alternation. 
%
Most SAT solvers can not efficiently handle the quantifier in the formulas. 
% 
Although, it is possible to encode the universal condition as a SAT or SMT problem.
%
There is always a danger of a blow-up in the formula size, making it harder for the SAT solver to handle. 
%Which may blow up the formula size for the solver to solve.
% 
As an experiment, we have implemented a direct SMT version (encoding not presented) of the universal problem extending the current $\fsmt$ search encodings.
%
The basic idea is to conjunct the formula for each combination value of the quantified variable. 
%
The Table~\ref{tab:satqbf-graph}, shows the search results of this encoding for the universal condition with 4 nodes and 3-connected graphs. 
%
The connectivity constraint and connectedness constraints are tagged with subscript $U$, $\texttt{StabilityC}_{U}$, $\texttt{Connected}(k)_{U}$,  to distinguish from earlier presented encodings of existential condition.
%
The SMT extended version was not able to scale it to the desired results and either timed-out (T/O) or ran out of memory (M/O).
%
This can be due to the fact that directly translating a quantified formula into a quantifier-free by introducing a vector of variables for each boolean possibility \change{Increase the formula size exponentially.. the current statement is a repeat of the forth line. Improve.}will blow up the size of the formula and may not be solvable by a solver.
%
%
\subsubsection{$\fqbf$ search encoding of the universal condition}
We will describe several constraints that encode universal condition based on QBF formulation in this section.
%
To avoid cumbersome notation, we will not explicitly write the ranges
of the indexing in the constraints.
%
$i$ and $j$ will range over nodes, i.e., from $1$ to $\nu$.
%
$m$ will range over molecules, i.e., from $1$ to $\mu$.
%
$q$ will range over edges between two nodes, i.e., from $1$ to $\pi$.
%
